---
title: "Unemployment Insurance Claims"
description: |
  County-level unemployment insurance data are hard to come by from the state, at least not in an easy to use format. We scrape claims data every week when an updated report is released by the state and make the data available for download here. We also share some comparisons of statewide claims versus claims for specific regions, over time.  
author:
  - name: Ani Ruhil
    url: {}
date: 01-09-2021
output:
  distill::distill_article:
    self_contained: false
draft: false
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE, message = FALSE, warning = FALSE, cache = TRUE,
  fig.align = "center", fig.width = 10, fig.height = 10, dev = "svg", fig.retina = 3
  )
```

People usually start focusing on unemployment insurance (UI) claims in the midst of an economic crisis, such as the one generated by the COVID-19 pandemic. Claims data then become a vital indicator of the labor market because they are released weekly rather than monthly unemployment rates the Bureau of Labor Statistics releases with a sizable lag. For example, at the time of writing this post the most recently available county-level unemployment rate data are for November 2020, and even these are provisional estimates likely to be revised. 

Timeliness is thus an asset of UI claims data, as is the fact that unlike the unemployment rate UI claims are based on administrative reports and hence likely more accurate. That said, however, one should beware of equating UI claims and the unemployment rate. Why? For one, filing for unemployment insurance does not necessarily mean each claim will result in the receipt of unemployment benefits. The unemployment rate is calculated from monthly surveys of households, and in these surveys you are flagged as unemployed if you have not worked in a number of weeks but are seeking employment. You may not have applied for UI benefits while currently unemployed, maybe because your job is not covered by unemployment insurance. 

All that an initial claim reflects is a filing by an unemployed individual after a separation from an employer, requesting a determination of basic eligibility for the UI program. In fact, this is what the [Bureau of Labor says about Initial and Continued Claims](https://www.dol.gov/ui/data.pdf).

> Initial Claim: An initial claim is a claim filed by an unemployed individual after a separation from an employer. The claimant requests a determination of basic eligibility for the UI program. When an initial claim is filed with a state, certain programmatic
activities take place and these result in activity counts including the count of initial claims. The count of U.S. initial claims for unemployment insurance is a leading economic indicator because it is an indication of emerging labor market conditions in the country. However, these are weekly administrative data which are difficult to seasonally adjust, making the series subject to some volatility.

> Continued Claim: A person who has already filed an initial claim and who has experienced a week of unemployment then files a continued claim to claim benefits for that week of unemployment. On a weekly basis, continued claims are also referred to as
insured unemployment, as continued claims reflect a good approximation of the current number of insured unemployed workers filing for UI benefits. The count of U.S. continued weeks claimed is also a good indicator of labor market conditions. While continued claims are not a leading indicator (they roughly coincide with economic cycles at their peaks
and lag at cycle troughs), they provide confirming evidence of the direction of the U.S. economy.

In the charts that follow we have calculated the rate of weekly initial claims filed in a jurisdiction per 100,000 persons (this is the **Rate** on the vertical axis). Scanning these data shows the stark break from the usual pattern of UI claims activity once the pandemic hit. After initial massive spikes in claims we see a subsidence but claims activity has yet to drop to pre-pandemic levels. What is more, some regions have it worse, as does Appalachia for example. And even though the general pattern is the same for every region or county, some have it worse. 


```{r datain}
load("~/Documents/Data Hub/UI Claims Data/data/rdata/uioh20.RData")

library(tidycensus)
load_variables(2019, "acs5") -> myvars
get_acs("county", state = "39", variable = "B01003_001") -> popn

library(tidyverse)
popn %>%
  mutate(
    county = stringr::str_remove_all(NAME, " County, Ohio")
    ) -> popn

popn %>%
  select(county, estimate) -> popn_df
```

We start with a simple comparison of Appalachia versus the state of Ohio. 

```{r data-cleaning}
uioh20 %>%
  left_join(popn_df, by = "county") %>%
  group_by(week_ending, Appalachia) %>%
  summarise(
    wk_date = lubridate::ymd(week_ending),
    persons = sum(estimate, na.rm = TRUE),
    initial = sum(initial_claims, na.rm = TRUE),
    continued = sum(continued_claims, na.rm = TRUE)
  ) %>%
  mutate(
    rate_initial = (initial * 100000) / persons,
    rate_continued = (continued * 100000) / persons
  ) %>%
  group_by(week_ending) %>%
  mutate(
    state_persons = sum(persons),
    state_initial = sum(initial),
    state_continued = sum(continued),
    strate_initial = (state_initial * 100000) / state_persons,
    strate_continued = (state_continued * 100000) / state_persons
  ) %>%
  ungroup() -> mydf

mydf %>%
  select(c(3, 9:13)) %>%
  distinct(wk_date, .keep_all = TRUE) -> mydf_state

mydf %>%
  select(c(2:8)) %>%
  distinct(wk_date, Appalachia, .keep_all = TRUE) %>%
  group_by(wk_date) %>%
  pivot_wider(
    names_from = Appalachia,
    values_from = c(persons, initial, continued, rate_initial, rate_continued)
  ) -> mydf_app
  

# The Mayors' Partnership for Progress
uioh20 %>%
  left_join(popn_df, by = "county") %>%
  mutate(
    MPP = case_when(
      county %in% c(
        "Athens", "Gallia", "Highland", "Hocking", "Jackson", "Lawrence", 
        "Meigs", "Monroe", "Morgan", "Muskingum", "Noble", "Perry", "Pike",
        "Ross", "Scioto", "Tuscarawas", "Vinton", "Washington") ~ "Mayor's Partnership",
    TRUE ~ "Rest of Ohio"
    )
  ) %>% 
  group_by(week_ending, MPP) %>%
  summarise(
    wk_date = lubridate::ymd(week_ending),
    persons = sum(estimate, na.rm = TRUE),
    initial = sum(initial_claims, na.rm = TRUE),
    continued = sum(continued_claims, na.rm = TRUE)
  ) %>%
  mutate(
    rate_initial = (initial * 100000) / persons,
    rate_continued = (continued * 100000) / persons
  ) -> mydf_mpp

mydf_mpp %>%
  distinct() -> mydf_mpp2

uioh20 %>%
  left_join(popn_df, by = "county") %>%
  mutate(
    MPP = case_when(
      county %in% c(
        "Athens", "Gallia", "Highland", "Hocking", "Jackson", "Lawrence", 
        "Meigs", "Monroe", "Morgan", "Muskingum", "Noble", "Perry", "Pike",
        "Ross", "Scioto", "Tuscarawas", "Vinton", "Washington"
        ) ~ "Mayors' Partnership",
      TRUE ~ "Rest of Ohio"
    )
  ) %>% 
  group_by(week_ending, county, MPP) %>%
  summarise(
    wk_date = lubridate::ymd(week_ending),
    persons = sum(estimate, na.rm = TRUE),
    initial = sum(initial_claims, na.rm = TRUE),
    continued = sum(continued_claims, na.rm = TRUE)
  ) %>%
  mutate(
    rate_initial = (initial * 100000) / persons,
    rate_continued = (continued * 100000) / persons
  ) -> mydf_mpp3

```

```{r the-plots}
ggplot() +
  geom_col(
    data = mydf_state,
    aes(x = wk_date, y = strate_initial)) +
  ylim(c(0, 3000)) +
  scale_x_date(date_breaks = "4 months", date_labels = "%b %d") +
  scale_y_continuous(labels = scales::comma) +  
  hrbrthemes::theme_ipsum_rc() +
  labs(x = "Week Ending", y = "Rate", title = "Ohio") -> p1

ggplot() +
  geom_col(
    data = mydf_app,
    aes(x = wk_date, y = rate_initial_No),
    alpha = 0.75) +
  ylim(c(0, 3000)) +
  scale_x_date(date_breaks = "4 months", date_labels = "%b %d") +
  scale_y_continuous(labels = scales::comma) +  
  hrbrthemes::theme_ipsum_rc() +
  labs(x = "", y = "", title = "Non-Appalachian") -> p2

ggplot() +
  geom_col(
    data = mydf_app,
    aes(x = wk_date, y = rate_initial_Yes),
    alpha = 0.50) +
  ylim(c(0, 3000)) +
  scale_x_date(date_breaks = "4 months", date_labels = "%b %d") +
  scale_y_continuous(labels = scales::comma) +  
  hrbrthemes::theme_ipsum_rc() +
  labs(x = "", y = "", title = "Appalachian") -> p3

ggplot() +
  geom_col(
    data = mydf_mpp2,
    aes(x = wk_date, y = rate_initial)) +
  ylim(c(0, 3000)) +
  scale_x_date(date_breaks = "4 months", date_labels = "%b %d") +  
  scale_y_continuous(labels = scales::comma) +
  hrbrthemes::theme_ipsum_rc() +
  labs(x = "Week Ending", y = "Rate", title = "The Mayors' Partnership") -> p4

library(patchwork)
p1 / (p2 + p3) / p4
```


What about individual county-level trends? 

```{r county-trends, layout = "l-page", out.width = "50%", fig.height = 10}
mydf_mpp3 %>%
  filter(MPP == "Mayors' Partnership") %>%
  ggplot() +
  geom_bar(aes(
      x = wk_date,
      y = rate_initial),
      stat = "identity") +
  facet_wrap(~ county, scales = "free_y", ncol = 3) +
  hrbrthemes::theme_ipsum_rc() +
  labs(
    title = "Ohio Law initial claims, representing newly unemployed persons",
    subtitle = "(For weeks ending 2020-01-04 through 2021-01-01)",
    caption = "@aruhil",
    x = "Week Ending",
    y = "Number of Initial Claims Filed per 100,000 persons") +
  scale_x_date(date_breaks = "4 months", date_labels = "%b %d") +
  scale_y_continuous(labels = scales::comma)
```



```{r pwalk}
library(purrr)

uioh20 %>%
  ungroup() %>%
  nest(-county) %>%
  pwalk(
      ~write_csv(
        x = .y, 
        paste0("~/Documents/github/theAllianceBlog/downloads/uiclaims/", .x, ".csv")
        )
      )

#uioh20 %>%
#  group_by(county) %>%
#    group_map(
#      ~distinct(
#        .x,
#        initial_claims, continued_claims, week_ending, Appalachia,
#        .keep_all = TRUE
#        ),
#      .keep_all = TRUE) %>% 
#  walk(
#    ~.x %>%
#      write_csv(path = paste0(".x", ".csv"))
#    )

```

County-level weekly (a) initial and (b) continued unemployment insurance claims data are available for **download below**. These files are updated weekly and the most recent update spans the **`r min(mydf$wk_date)`** ... **`r max(mydf$wk_date)`** period. 


|   |   |   |   |   |   |   |   |   |   |   |
|:--|:--|:--|:--|:--|:--|:--|:--|:--|:--|:--|
| [Adams](downloads/uiclaims/Adams.csv)   | [Allen](downloads/uiclaims/Allen.csv)   | [Ashland](downloads/uiclaims/Ashland.csv)   | [Ashtabula](downloads/uiclaims/Ashtabula.csv)   |  [Athens](downloads/uiclaims/Athens.csv)    |  [Auglaize](downloads/uiclaims/Auglaize.csv)    |   [Belmont](downloads/uiclaims/Belmont.csv)   | [Brown](downloads/uiclaims/Brown.csv)   | [Butler](downloads/uiclaims/Butler.csv)    | [Carroll](downloads/uiclaims/Carroll.csv)   | [Champaign](downloads/uiclaims/Champaign.csv)   |
| [Clark](downloads/uiclaims/Clark.csv)   | [Clermont](downloads/uiclaims/Clermont.csv)   |   [Clinton](downloads/uiclaims/Clinton.csv)   |  [Columbiana](downloads/uiclaims/Columbiana.csv)   |  [Coshocton](downloads/uiclaims/Coshocton.csv)    |  [Crawford](downloads/uiclaims/Crawford.csv)    |  [Cuyahoga](downloads/uiclaims/Cuyahoga.csv)   | [Darke](downloads/uiclaims/Darke.csv)   | [Defiance](downloads/uiclaims/Defiance.csv)    | [Delaware](downloads/uiclaims/Delaware.csv) | [Erie](downloads/uiclaims/Erie.csv)   |
| [Fairfield](downloads/uiclaims/Fairfield.csv) | [Fayette](downloads/uiclaims/Fayette.csv)   | [Franklin](downloads/uiclaims/Franklin.csv)   | [Fulton](downloads/uiclaims/Fulton.csv)   | [Gallia](downloads/uiclaims/Gallia.csv)   | [Geauga](downloads/uiclaims/Geauga.csv)    | [Greene](downloads/uiclaims/Greene.csv)    | [Guernsey](downloads/uiclaims/Guernsey.csv)   |   [Hamilton](downloads/uiclaims/Hamilton.csv)   |   [Hancock](downloads/uiclaims/Hancock.csv)    | [Hardin](downloads/uiclaims/Hardin.csv)   |
| [Harrison](downloads/uiclaims/Harrison.csv)   | [Henry](downloads/uiclaims/Henry.csv)   | [Highland](downloads/uiclaims/Highland.csv)   |  [Hocking](downloads/uiclaims/Hocking.csv)   | [Holmes](downloads/uiclaims/Holmes.csv)   | [Huron](downloads/uiclaims/Huron.csv)    | [Jackson](downloads/uiclaims/Jackson.csv)    | [Jefferson](downloads/uiclaims/Jefferson.csv)   | [Knox](downloads/uiclaims/Knox.csv)   |  [Lake](downloads/uiclaims/Lake.csv)    | [Lawrence](downloads/uiclaims/Lawrence.csv)   |
| [Licking](downloads/uiclaims/Licking.csv)   | [Logan](downloads/uiclaims/Logan.csv)   |   [Lorain](downloads/uiclaims/Lorain.csv)   |   [Lucas](downloads/uiclaims/Lucas.csv)   |   [Madison](downloads/uiclaims/Madison.csv)   |   [Mahoning](downloads/uiclaims/Mahoning.csv)    |   [Marion](downloads/uiclaims/Marion.csv)    |   [Medina](downloads/uiclaims/Medina.csv)   |   [Meigs](downloads/uiclaims/Meigs.csv)   |   [Mercer](downloads/uiclaims/Mercer.csv)    |   [Miami](downloads/uiclaims/Miami.csv)   |
| [Monroe](downloads/uiclaims/Monroe.csv) | [Montgomery](downloads/uiclaims/Montgomery.csv)   |  [Morgan](downloads/uiclaims/Morgan.csv)   |   [Morrow](downloads/uiclaims/Morrow.csv)   |   [Muskingum](downloads/uiclaims/Muskingum.csv)   |   [Noble](downloads/uiclaims/Noble.csv)    |   [Ottawa](downloads/uiclaims/Ottawa.csv)    |   [Paulding](downloads/uiclaims/Paulding.csv)   |   [Perry](downloads/uiclaims/Perry.csv)   |   [Pickaway](downloads/uiclaims/Pickaway.csv)    |   [Pike](downloads/uiclaims/Pike.csv) |
| [Portage](downloads/uiclaims/Portage.csv) | [Preble](downloads/uiclaims/Preble.csv)   |   [Putnam](downloads/uiclaims/Putnam.csv)   |   [Richland](downloads/uiclaims/Richland.csv)   |   [Ross](downloads/uiclaims/Ross.csv)   |   [Sandusky](downloads/uiclaims/Sandusky.csv)    |   [Scioto](downloads/uiclaims/Scioto.csv)    |   [Seneca](downloads/uiclaims/Seneca.csv)   |   [Shelby](downloads/uiclaims/Shelby.csv)   |   [Stark](downloads/uiclaims/Stark.csv)    |   [Summit](downloads/uiclaims/Summit.csv)   |
| [Trumbull](downloads/uiclaims/Trumbull.csv) | [Tuscarawas](downloads/uiclaims/Tuscarawas.csv)   | [Union](downloads/uiclaims/Union.csv)   |   [Van Wert](downloads/uiclaims/VanWert.csv)   |   [Vinton](downloads/uiclaims/Vinton.csv)   |   [Warren](downloads/uiclaims/Warren.csv)    |   [Washington](downloads/uiclaims/Washington.csv)    |   [Wayne](downloads/uiclaims/Wayne.csv)   |   [Williams](downloads/uiclaims/Williams.csv)   |   [Wood](downloads/uiclaims/Wood.csv)    |   [Wyandot](downloads/uiclaims/Wyandot.csv) |






